---
title: "CSLS_weather"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CSLS_weather}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Weather

Quick way of visualizing wet/dry years over long time and short time in the
Central Sands. Uses weather data from Hancock, WI, so this is best for
visualization purposes - it's not exact match for the CSLS MODFLOW model. Colors
represent annual precipitation minus potential evapotranspiration as calculated
using the Hamon method. Blue is wetter years, red is drier years. See actual
data for exact numbers - again, this is meant to be a quick, intuitive about how
wet years intermix with dry years.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(CSLSevap)
library(NISTunits)
library(lubridate)
library(zoo)
library(ggplot2)
library(extrafont)

# May need to change this line based on your working directory. Needs to be
# relative to vignetted dir when building vignettes, but from root dir when
# checking and building package.
# hancock      <- read.csv("../data-raw/hancock_ppt_19030101_20210305.csv") 
hancock      <- read.csv("hancock_ppt_19030101_20210305.csv") 
hancock      <- hancock %>%
                mutate(date = as_datetime(.data$DATE),
                       phi = NISTdegTOradian(.data$LATITUDE),
                       pcpn = NISTinchTOmeter(.data$PRCP)*1000,
                       tmin = NISTdegFtOdegC(.data$TMIN),
                       tmax = NISTdegFtOdegC(.data$TMAX)) %>%
                dplyr::select(.data$phi, .data$date, .data$pcpn, .data$tmin, .data$tmax)
loc          <- list(phi = hancock$phi)
dates        <- data.frame(date = seq(min(hancock$date), max(hancock$date), "1 day"))
hancock      <- left_join(dates, hancock, by = "date") %>%
                dplyr::select(.data$date, .data$pcpn, .data$tmin, .data$tmax)
hancock.zoo  <- read.zoo(hancock)
hancock.zoo  <- na.approx(hancock.zoo)
hancock$tmin <- as.numeric(hancock.zoo$tmin)
hancock$tmax <- as.numeric(hancock.zoo$tmax)
hancock$pcpn[is.na(hancock$pcpn)] <- 0
  

weather <- list(datetimes = hancock$date,
                atmp = list(min = hancock$tmin,
                            max = hancock$tmax))

evap <- evaporation(method = "Hamon", loc, weather)
hancock$pet <- evap


annual_water <- hancock %>%
                mutate(pcpn_pet = .data$pcpn - .data$pet) %>%
                group_by(year = year(.data$date)) %>%
                summarise(pcpn = sum(.data$pcpn),
                          pet = sum(.data$pet),
                          pcpn_pet = sum(.data$pcpn_pet),
                          .groups = "drop") %>%
                mutate(y = 0)

```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=6.5}

plot_color_strip <- function(df, 
                             year_start = NULL,
                             year_end = NULL, 
                             text_size = 10,
                             legend_pos = "left") {
  
  if (!is.null(year_start)) {
    df <- df %>% filter(.data$year >= year_start)
  }
  if (!is.null(year_end)) {
    df <- df %>% filter(.data$year <= year_end)
  }
  
  plot_obj <- ggplot(df) +
              geom_tile(aes(x = .data$year,
                            y = .data$y,
                            fill = .data$pcpn_pet),
                        color = NA) +
              scale_fill_distiller(name = "",
                                   palette = "RdBu",
                                   direction = 1) +
                                   # limits = c(-1,1)*max(abs(df$pcpn_pet)))+
              scale_y_continuous(expand = c(0,0)) +
              scale_x_continuous(expand = expansion(add = c(1,1))) +
              labs(x = "") +
              theme_classic() +
              theme(axis.text.y = element_blank(),
                    axis.title.y = element_blank(),
                    axis.line.y = element_blank(),
                    axis.ticks.y = element_blank(),
                    text = element_text(family = "Segoe UI Semilight",
                                        size = text_size),
                    legend.position = legend_pos,
                    legend.text = element_blank())
  return(plot_obj)
  
}


```


### 1938 to 2020 Timeseries

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=7.5}
p1 <- plot_color_strip(annual_water, 1938, 2020, text_size = 16,
                       legend_pos = "top")
p1

ggsave("color_strip_1938.png", p1, device = "png", 
       width = 7.5, height = 2, units = "in")

```


### 1981 to 2020 Timeseries

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=1.6, fig.width=4}
rejigger_years <- annual_water %>% mutate(year = .data$year - 1981+1)
p2 <- plot_color_strip(rejigger_years, 1, 38, text_size = 16)
p2
ggsave("color_strip_1981.png", p2, device = "png", 
       width = 4, height = 1.6, units = "in")
```
